---
title: "Homework 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
```

## Problem \#1: Testing coin flips <small>(6 pts)</small>

In the six sequences below, only one of them is actually **randomly generated from independent flips of a fair coin**. Use a combination of everything you know (common sense, Monte Carlo, hypothesis testing, etc.) to identify which is actually random and explain your reasoning.

(For full points, conduct a formal test and report a $p$-value for each sequence. You may use a combination of multiple tests to arrive at your answer. If you cannot compute a p-value for each sequence, you can still earn a significant amount of partial credit by carefully explaining your reasoning and response as best as you can.)

My advice is **be creative** with the test statistics you come up with to eliminate each sequence! Think of some way of summarizing a sequence of flips that might be useful for comparing against a simulated sequence of random flips. After you come up with an idea for a statistic, remember to run it on many MC generated completely random flips to produce a distribution under the null, which you can then compare with your data to get a p-value. Also, be careful of now you define "more extreme" than the data.

(2 bonus points available if you can find a single test that is powerful enough to reject all the fake sequences together in one step. Yes, at least one such possible test exists.)

```{r}
flips1 = "HTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHT"

flips2 = "HHHTHTTTHHTHHTHHHTTTTHTHTHHTTHTHHHTHHTHTTTHTHHHTHTTTHTHTHHTHTHTTHTHHTHTHTTTHTHHHTHTHTTHTHTHHTHTHTHHHTHTTTHTHHTHTHTHHTTTHTHHTHHTTTTHTHTHHHTHTTHTHHTHTHTTHTHHTHTHHHTHHHTHTTTHTTHTTTHTHHHTHTHTTHTHHTHHTHTTT"

flips3 = "HHTHTHTTTHTHHHTHHTTTHTHHTHTTTHTHTHHTHTHTTHTHHHHHHTTTHTHTHHTHTTTHTHHTHTHTTTHTHHHTTHTTTHTHTHHHHTHTTHHTTTTTHTHHHTHTHTTTTTHHHTHHTHHTHHHTTTTHTHTHHHTHHTTTTTHTHHHTHTHTHTTTHTHHHTHTHTHTTHTHHTHTHTHTTTTHTHHHTHTH"

flips4 = "HTHHHHHHHTHTTHHTTHHHTHTHTTTHHTHHHTHHTTHTTTTTTTTTHTHHTTTTTHTHTHTHHTTHTTHTTTTTHHHTHTTTHTHTHHHTHTTTTHTHTHHTTHTHTTHHTHTHHHHTHTTHHTTHTTHTTHTHHHHHHTTTTTTHHHTTHTHHHHTTTHTTHHHTTHTHHTTTHHTHHTTTHTHHTHHHTHHTTHHH"

flips5 = "HHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTT"

flips6 = "TTHTTTHTTTTTTTHTHTHTHTTHTTHTHHTHHTTTHHTHTTTHTHHTHHHTHTTHHTHHTTHTHTTTTHTHTTTHHTTTTTTTTHTHHTTHTTTTTTHTHTHTHTTTHTTHHTTHTTTHHTTTHTTHTTTTHTTTTHHTTTHTHTHHHTTTTTTHTHHTTTTTTTTTTTTHHHTTTHHHTTTHTTTHTHTTHTTTTTHT"

# you can use the function below to split the above sequences in vectors of flips
split = function(str) strsplit(str, split="")[[1]]
split(flips1)
```



## Problem \#2: Finding the Trick Coin <small>(6 pts; 2pts each)</small>

I have two coins in my pocket - a trick coin with two heads and a fair coin with one head and one tail(s?). We'll play a game. I will grab one coin at random, and flip it $N$ times. After that you will have to decide if it is the fair coin or the trick coin. The null hypothesis is that it is the fair coin. 

**Decision Rule 1**: If after $N$ flips there are no tails, then you decide it is the trick coin. If there is at least 1 tail then you know it is the fair coin. 

a. Using "Decision Rule 1", what is the lowest number of flips $N$ would you need in order to have a significance level less than 5% for this test?
b. Using $N$ from part a, what is the power of the test?
c. Suppose $N=4$ is decided. How can you modify the decision process to have a significance level of exactly 5%? Does this change the power of the test?
d. Extra Credit (2 points): Suppose if you guess correct you win \$100 (and if you're wrong you get nothing), but each flip of the coin costs \$10. What strategy would you use to maximize your expected profit from this game?


## Problem \#3: Testing the maximum of a uniform distribution <small>(8 pts; 2 pts each)</small>

We sample $X_1, X_x,\ldots,X_n \overset{\text{iid}}\sim\text{Uniform}(0,m)$ where $m$ is an unknown maximum. Sleazy Jim tells you that $m=1$ but you're not so sure. The 50 values sampled are in the following data file:

```{r}
X <- read.csv("uniform_sample.csv")$x
```

a. Write out in formal notation the null and alternative hypotheses. 
b. Come up with a test statistic and measure your sampled data. Is this a one-sided test or two-sided test?
c. Simulate a distribution for the test statistic under the null hypothesis of size at least 1000. Display a histogram of your test statistic distribution.
d. Calculate the $p$-value for this data and make a conclusion.

## Problem \#4: Blurtle <small>(10 pts; 2 pt each)</small>

Have you been playing the hot new game Blurtle? It's a (fictional) word game you can play daily - you have to guess a 5 letter word and you only have 6 attempts. I've been playing for the past 100 days and I've been tracking my number of guesses. I'm trying to figure out whether I have been getting better or not.

The file `blurtle.csv` contains 100 rows of data, giving the number of tries to guess the word. If it took 7 guesses that actually means I failed (you don't actually get a 7th guess). 

```{r}

# Load necessary libraries
library(dplyr)

# Read the data from the CSV file
blurtle <- read.csv("blurtle.csv")
blurtle
blurtle$tries
mean(blurtle$tries)
var(blurtle$tries)

# Compute observed test statistic
observed_stat <- cor(blurtle$day, blurtle$tries, method = "spearman")

# Permutation test function
perm_test <- function(data, n_perm = 1000) {
  perm_stats <- numeric(n_perm)
  
  for (i in 1:n_perm) {
    perm_data <- data
    perm_data$tries <- sample(perm_data$tries)
    perm_stats[i] <- cor(perm_data$day, perm_data$tries, method = "spearman")
  }
  
  return(perm_stats)
}

set.seed(123)
perm_stats <- perm_test(blurtle)

# Calculate p-value
p_value <- (sum(abs(perm_stats) >= abs(observed_stat)) + 1) / (length(perm_stats) + 1)

# Plot distribution of permutation test statistics
hist(perm_stats, main = "Permutation Test Statistics", xlab = "Spearman Correlation", breaks = 30)
abline(v = observed_stat, col = "red", lwd = 2)

# Print results
print(paste("Observed test statistic: ", observed_stat))
print(paste("P-value: ", p_value))



```

Your task is to perform a permutation test on the data to determine if there is statistical evidence of a true improvement trend.

a. State the null and alternative hypotheses
H0: There is no change in the trend in the number of tries.
H1: There is change in the trend in the number of tries

b. Determine a test statistic that identifies a trend in the number of tries. There are many good ones you could use - be creative.

Spearman rank correlation coefficient between day and number of tries can be used. This statistic will capture the monotonic trend, indicating if there’s a general improvement or deterioration over time.

c. Decide whether the test will be a one or two-tailed test

Two-tailed test is used because we are looking for either improvement or setback  - lower or greater number of guesses

d. Simulate a distribution of test statistics under the null hypothesis

usage of perturbation test is implementd

e. Calculate the test statistic on the observed data, calculate the $p$-value and state your conclusions.

 "Observed test statistic:  -0.230958875668475"
[1] "P-value:  0.023976023976024"

We conclude that there is enough evidence to reject the H0 hypothesis. Test statistic has negative value meaning that there is enough evidence that number of guesses by days is lowering.


II. version of solution
Linear Regression Slope: Fit a simple linear regression model with day as the predictor and tries as the response variable. The slope of the regression line will indicate the trend. A significantly negative slope would indicate improvement.

```{r}

# Load necessary library
library(dplyr)

# Read the data from the CSV file
blurtle <- read.csv("blurtle.csv")

# Compute the linear regression slope
model <- lm(tries ~ day, data = blurtle)
slope <- coef(model)[2]

# Permutation test function for the slope
perm_test_slope <- function(data, n_perm = 1000) {
  perm_slopes <- numeric(n_perm)
  
  for (i in 1:n_perm) {
    perm_data <- data
    perm_data$tries <- sample(perm_data$tries)
    perm_model <- lm(tries ~ day, data = perm_data)
    perm_slopes[i] <- coef(perm_model)[2]
  }
  
  return(perm_slopes)
}

set.seed(123)
perm_slopes <- perm_test_slope(blurtle)

# Calculate p-value
p_value_slope <- (sum(abs(perm_slopes) >= abs(slope)) + 1) / (length(perm_slopes) + 1)

# Plot distribution of permutation test slopes
hist(perm_slopes, main = "Permutation Test Slopes", xlab = "Slope", breaks = 30)
abline(v = slope, col = "red", lwd = 2)

# Print results
print(paste("Observed slope: ", slope))
print(paste("P-value: ", p_value_slope))


```
the same conclusion. Negative slope meaning that with days number of guesses is lover, confirmed by p value.

in this solution we use runs.test. The result will indicate whether there is statistical evidence of a trend (either improving or worsening).

```{r}
# Load necessary library
library(tseries)

# Read the data from the CSV file
blurtle <- read.csv("blurtle.csv")
head(blurtle)
# Calculate changes from the previous day (1 for improvement, 0 for no improvement or worsening)
improvement <- ifelse(diff(blurtle$tries) < 0, 1, 0)
improvement

# Perform the runs test
runs_test_result <- runs.test(as.factor(improvement))

# Print results
print(runs_test_result)
```

# of guesses per day.  45 means 0 improvement, 55 0 improvement, 53  1 improvement 43 0 improvement....
[1] 4 5 5 3 4 3 3 5 5 4 4 4 5 4 6 5 5 7 4 3 6 4 5 4 4 4 4 4 6 4 5 5 5 4 3 5 3 4 5 4 3 5 3 5 5 3 4 5 5
[50] 5 4 3 4 4 4 4 4 5 5 4 6 4 4 4 5 4 4 3 4 3 4 3 4 4 4 6 3 4 3 2 4 5 4 4 3 4 5 4 4 4 4 4 5 3 3 3 4 4

p= 0.006313  ->  sequence is not random, and we count   45  not improvement and  43 improvement  z=2.73  is right side




```{r}
# Load necessary libraries
library(ggplot2)

# Load the blurtle.csv file
blurtle <- read.csv("blurtle.csv")

# Compute the observed Spearman correlation
observed_stat <- cor(blurtle$day, blurtle$tries, method = "spearman")
print(paste("Observed Spearman correlation: ", observed_stat))

# Permutation test function
perm_test <- function(data, n_perm = 1000) {
  perm_stats <- numeric(n_perm)
  
  for (i in 1:n_perm) {
    # Permute the 'tries' column (shuffle it randomly)
    perm_data <- data
    perm_data$tries <- sample(perm_data$tries)
    
    # Compute the Spearman correlation for the permuted data
    perm_stats[i] <- cor(perm_data$day, perm_data$tries, method = "spearman")
  }
  
  return(perm_stats)
}

# Set a seed for reproducibility
set.seed(123)

# Run permutation test (1000 permutations)
perm_stats <- perm_test(blurtle, n_perm = 1000)

# Plot the distribution of permuted test statistics
hist(perm_stats, main = "Permutation Test Statistics", 
     xlab = "Spearman Correlation", breaks = 30, col = "lightblue")
abline(v = observed_stat, col = "red", lwd = 2)

# Calculate p-value
p_value <- (sum(abs(perm_stats) >= abs(observed_stat)) + 1) / (length(perm_stats) + 1)
print(paste("P-value: ", p_value))

# Interpretation
if (p_value < 0.05) {
  print("There is enough evidence to reject the null hypothesis. The number of tries shows a significant trend over time.")
} else {
  print("There is not enough evidence to reject the null hypothesis. The number of tries does not show a significant trend over time.")
}

```
```{r}
# Load necessary libraries
library(ggplot2)

# Scatter plot of 'day' vs 'tries'
ggplot(blurtle, aes(x = day, y = tries)) +
  geom_point(color = "blue", size = 2) +            # Scatter points
  labs(title = "Blurtle: Number of Tries Over Time", 
       x = "Day", 
       y = "Number of Tries") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # Add a linear trend line
  theme_minimal()

```

