---
title: "Homework 11"
author: "Matej Popovski"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
```

## Problem 1. Warm up: Log-Odds, Odds and Probability <small>5 points; 1 point each</small>

### Part a) 
Convert a log-odds of 1.25 to probability.

***

odds = p / (1 - p)
odds*(1-p) = p
odds - p*odds = p
odds = p(1 + odds)
p = odds / (1 + odds)

log-odds = 1.25
odds = e^1.25

p = e^1.25 / (1 + e^1.25)

p = 0.77729986117

***

### Part b) 
Convert probability of 0.65 to odds.

***

odds = p / (1 - p)
odds = 0.65 / 0.35 = 1.857143

***

### Part c)
Event $A$ has a probability of 0.8 and the odds ratio of $A$ to $B$ is 1.3. Calculate the probability of event $B$.

***
odds(A)/odds(B) = 1.3

odds(A) = 0.8 / 0.2 = 4

odds(B) = odds(A) / 1.3 = 4 / 1.3 = 3.076923

p(B) = odds(B) / (1 + odds(B)) = 3.076923 / (1 + 3.076923)

p(B) = 0.754717

***

### Part d) 
You roll a single 6 sided die. What are the odds you get a number divisible by 3?


***

p(divisible by 3) = 1/3

odds = 1/3 / 2/3 = 1/2

***

### Part e) 
The odds ratio comparing event A to B is 9 and the risk ratio is 3. What is $Pr(A)$? *Hint: let $Pr(A)=a$ and $Pr(B)=b$, and write out the odds ratio and risk ratio as equations. 

***

rr = p(A) / p(B) = 3 

p(A) = 3*p(B)

oddsR = odds(A) / odds(B) = 9

p(A)/(1 - p(A)) / p(B)/(1 - p(B)) = 9

3p(B)/(1-3p(B)) / p(B)/(1 - p(B)) = 9

3 * (1 - p(B)) /(1-3p(B)) = 9

(1 - p(B)) /(1-3p(B)) = 3

1 - p(B) = 3 - 9p(B)

2 = 8p(B)

p(B) = 1/4

p(A) = 3/4

***


## Problem 2. Interpreting logistic regression <small>6pts; 2 pts each</small>

Suppose we collect data for a group of students in a statistics class with independent variables $X_{1}=\text{hours studied}$, $X_{2}=\text{GPA}$, and binary response variable
$$
Y= \begin{cases} 1 &\mbox{ if student received an A} \\
  0 &\mbox{ otherwise. }
  \end{cases}
$$
Suppose that we fit a logistic regression model to the data, predicting $Y$ from $X_1$ and $X_2$ (and an intercept term) and produce estimated coefficients $\hat{\beta}_{0}=-5, \hat{\beta}_{1}=0.06, \hat{\beta}_{2}=1.1$.

### Part a) Logistic regression and probability
According to our fitted model, what is the probability that a student receives an A if they study for $40$ hours and have a GPA of $3.5$?

> sigma = exp{beta0 + beta1x1 + beta2x2} / (1 + exp{beta0 + beta1x1 + beta2x2})

sigma = exp{-5 + 0.06 40 + 1.1 3.5} / (1 + exp{-5 + 0.06 40 + 1.1 3.5}) = 0.7775

### Part b) Interpreting coefficients
According to our fitted model, an additional hour spent studying is associated with *how much* of an increase in the log odds of receiving an A?

An additional hour in studying is associated with 0.06 increase in the log-odds of receiving an A, holding all other variables constant.

### Part c) "Inverting" logistic regression probabilities
According to our fitted model, how many hours would the student in Part (a) need to study to have a $50\%$ chance of getting an A in the class?
That is, keeping GPA fixed at $3.5$, how many hours of study are needed so that the predicted probability of an A is $50\%$?
If you aren't up for the math, feel free to find an approximate solution via guess-and-check in R.


## Problem 3. Palmer Penguins Part 1  <small>9pts; 3 pts each</small>

The Palmer Penguin dataset (https://allisonhorst.github.io/palmerpenguins/) consists of 344 observations of penguins belonging to 3 penguin species across the islands in the Palmer Archipeligo in Antarctica. We will build a logistic model attempting to classify the penguins based on physical characteristics. For each penguin we've recorded: 

* `species` - the species, either "Gentoo", "Adelie" or "Chinstrap"
* `island` - which of three islands the Penguin was observed (Biscoe, Dream or Torgersen)
* `bill_length_mm` - the length of the bill in mm
* `bill_depth_mm` - the depth of the bill (vertical thickness of the closed bill)
* `filler_length_mm` - length of their cute flippers
* `body_mass_g` - the body mass in grams
* `sex` - female, male or NA (unknown)
* `year` - The year of the observation: 2007, 2008 or 2009

First you need to download the library. Run this chunk of code once.
```{r, eval=FALSE, echo=FALSE}
#Run this code once to install the library
install.packages("palmerpenguins")
```

Then load the library and the penguin dataset. Note: Your RMD won't knit until you run the above chunk.
```{r}
library(palmerpenguins)
```

### a) Adelie penguins based on island

We are going to try to classify penguins as **Adelie** or **not Adelie**. So create a new variable called `Adelie` which will be 1 or 0 based on whether the penguin species is Adelie

Perform some analysis looking at each of the 3 islands - create a 2 way table between `island` and the `Adelie` variable. Look at the proportions conditioned on island. What proportion of observations on each island were Adelie?


### b) Adelie Penguins on Dream

Find the (i) probability, (ii) odds and (iii) log-odds that a randomly selected penguin from Dream is an Adelie penguin?


### c) An island-based classifier

Now fit a logistic model predicting whether a penguin is Adelie based on island.
Interpret the intercept and the coefficient of the `islandDream` variable. Use this model to predict the probability that a penguin from Dream is Adelie.




## Problem 4. Penguins Part II <small>10pts; 2 pts each</small>

In this problem we will work once again with the Palmer Penguin dataset. We will work with a subset by taking out all missing values. After you have installed the package and loaded the library, uncomment the line below.
```{r}
 penguins.complete <- penguins[complete.cases(penguins),]
```

### a) Predicting Palmer Penguins with quantitative Predictors

Now use the two bill measurements (`bill_length_mm` and `bill_depth_mm`) as predictors in a new logistic model. Suppose a penguin with a bill length of 54.1 and a bill depth of 23.7 is observed. What is the model's probability that the penguin is an Adelie penguin? *Hint: Your response variable in the formula can be `species=="Adelie"` and you won't have to create a new binary response variable column.*


### b) Interpreting coefficients

Are longer bills associated with an increased or decreased likelihood that a penguin is an Adelie penguin?


### c) A full classifier

Fit a logistic model to predict whether a penguin is a **Chinstrap** penguin using all four of the biological measurements (`bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`). 


Which of the predictors are significant?



### d) Assessing the model
Suppose you will predict that a penguin is a Chinstrap if the estimated $\hat{y}=\sigma(\hat{z})>0.5$. When predicting using this threshold, what is the type 1 error rate of your predictor on the dataset? What is power of the predictor on the dataset? 
*Hint: you will want to compare the predicted $\hat{y}$ values to the actual $y$ values. The `table` command can produce a 2x2 confusion matrix to help you answer this question.*




### e) Adjusting the Type 1 error rate

Now modify your threshold from 0.5 to some other threshold with the goal of achieving the highest power possible while keeping the type 1 error rate  below 0.05. What threshold would you use? What is the type 1 error rate and power of this new classifier?


