
---
title: "STAT340  Estimation Examples"
author: "Brian Powers"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

## Experiments with Estimator Bias

Let's consider $S^2$, the sample variance.

Let's consider a population that is normally distributed, say $N(100, 5^2)$, so $\sigma=5$. Let's look at samples of size 20.

```{r}
sample_vars <- 0
NMC <- 10000
for(i in 1:NMC){
  mysample <- rnorm(20, 100, 5)
  sample_vars[i] <- var(mysample)
}
hist(sample_vars)
abline(v=mean(sample_vars), lwd=3)
abline(v=5^2, col="red", lty=2, lwd=3)
```

Just for kicks, let's repeat with a Poisson population with parameter $\lambda=5$, still a sample size of $n=20$
```{r}
sample_vars <- 0
NMC <- 10000
for(i in 1:NMC){
  mysample <- rpois(20, lambda=5)
  sample_vars[i] <- var(mysample)
}
hist(sample_vars)
abline(v=mean(sample_vars), lwd=3)
abline(v=5, col="red", lty=2, lwd=3)


```

We can use the same simulation to look at the bias of the sample standard deviation (back to the normal distribution)

```{r}
NMC <- 10000
sample_sds <- 0 #empty vector
for(i in 1:NMC){
  mysample <- rnorm(20, 100, 5)
  sample_sds[i] <- sd(mysample)
}
hist(sample_sds, main="Sampling distribution of Sample SD")
abline(v=mean(sample_sds), lwd=3)
abline(v=5, col="red", lty=2, lwd=3)
```

We might be interested in the bias as it relates to sample size.
Let's look at samples of size 5, 10, 15, ..., 100

```{r}
ns <- seq(5, 100, 5)
biases <- rep(0, length(ns))
for(i in 1:length(ns)){
NMC <- 10000
  sample_sds <- 0
  for(j in 1:NMC){
    mysample <- rnorm(ns[i], 100, 5)
    sample_sds[j] <- sd(mysample)
  }
  biases[i] <- mean(sample_sds) - 5
}
plot(x=ns, y=biases, type="l", main="bias of s as n increases", ylim=c(min(biases),0))
abline(h=0, lty=2)
```



Let's look at the sample maximum as an estimate of the population maximum.
Let's sample data from a uniform(0, 1), and we'll look at a sample of size 20.
```{r}
NMC <- 10000
sample_maxes <- 0
for(i in 1:NMC){
  mysample <- runif(20, 0, 1)
  sample_maxes[i] <- max(mysample)
}
hist(sample_maxes, main="Sampling distribution of Sample Max", breaks=50)
abline(v=mean(sample_maxes), lwd=3)
abline(v=1, col="red", lty=2, lwd=3)
```

It turns out that we can modify the estimator to correct for the bias.
If we use $\dfrac{n+1}{n}X_{(n)}$ this gives us an unbiased estimator for population maximum (you can look this up if you don't believe me). Let's demonstrate:
```{r}
NMC <- 10000
n <- 20
sample_maxes_adj <- 0
for(i in 1:NMC){
  mysample <- runif(n, 0, 1)
  sample_maxes_adj[i] <- (n+1)/(n)*max(mysample)
}
hist(sample_maxes_adj, main="Sampling distribution of adjusted Sample Max", breaks=40)
abline(v=mean(sample_maxes_adj), lwd=3)
abline(v=1, col="red", lty=2, lwd=3)
```

### German Tank Example

Suppose we have tanks produced each with a serial number from 1 up to N. We are able to capture and observe the serial numbers on a sample of say 5 tanks. Can we estimate the population size of tanks (get an estimate for N)???

By the same logic from above, one might guess that (n+1)/n * max(X) would be a good estimate of N. Let's imagine that there are 100 tanks in the population.
```{r}
NMC <- 10000
N <- 100
n <- 5
sample_maxes <- 0
for(i in 1:NMC){
  mysample <- sample(1:N, size=n, replace=FALSE)
  sample_maxes[i] <- (n+1)/(n)*max(mysample)
}
hist(sample_maxes, main="Sampling distribution of adjusted Sample Max")
abline(v=mean(sample_maxes), lwd=3)
abline(v=N, col="red", lty=2, lwd=3)
```
For the discrete case we actually need to use $(n+1)/n*max(X) - 1$ as our unbiased estimator. The reason is that the discrete uniform behaves slightly differently than the continuous - because the probability of observing the population max is > 0.

```{r}
NMC <- 10000
N <- 100
n <- 5
sample_maxes <- 0
for(i in 1:NMC){
  mysample <- sample(1:N, size=n, replace=FALSE)
  sample_maxes[i] <- (n+1)/(n)*max(mysample)-1
}
hist(sample_maxes, main="Sampling distribution of adjusted Sample Max")
abline(v=mean(sample_maxes), lwd=3)
abline(v=N, col="red", lty=2, lwd=3)
```



### Capture - Recapture population estimation

Suppose that a population is of size N. One day you capture and tag a random sample of $n_1$ individuals and release them back into the wild.
Several weeks later you capture a new batch of size $n_2$ and $k$ of the new sample are tagged. How can we estimate the population size from this?

Suppose the population is of size 400 and we tag and release 25. Suppose our second sample is of size 20. How many would we expect to see in the second sample?

```{r}
N <- 30000
n1 <- 450
n2 <- 500
population <- c(rep(1, n1), rep(0, N-n1))

NMC <- 10000
recaptured <- replicate(NMC, 
                        sum(sample(population, n2)))
barplot(table(recaptured))
```
What would be a good estimate of the population size? 

After the first tagging, we know that the population has exactly $n_1$ tagged individuals. So assuming (1) that the population has been randomized and (2) that deaths are not consequential (either not many, or that both tagged/untagged were equally likely to have died) then the probability of a tagged fish being recaptured is $n_1/N$. So the proportion of tagged fish in the second sample, which is $k/n_2$ should be the same as $n_1/N$. 

$$\frac{k}{n_2} = {n_1}{N}$$

Under this logic, we could solve for $N$ to find a point estimate

$$\hat{N}=\frac{n_1 n_2}{k}$$

For example, if we found 7 in the second sample we'd calculate
```{r}
450*500/7
```

It seems to be an over-estimate. We can examine the bias by repeating the simulation
Note - we have to throw out any recaptures where 0 are tagged - obviously infinity is not a useful estimate.

```{r}
estimates <- n1*n2/recaptured
hist(estimates, breaks=40)
abline(v=mean(estimates[estimates<Inf]), lwd=3)
abline(v=N, col="red", lty=2, lwd=3)

```




### Example: Estimating the rate parameter in the exponential distribution.

```{r}
run_exprate_expt <- function( n, rate ) {
  data <- rexp(n=n, rate=rate);
  Xbar <- mean( data )
  return( 1/Xbar )
}
```

```{r}
M <- 1e4;
replicates <- rep(NA, M);
for (i in 1:M ) {
  replicates[i] <- run_exprate_expt(2000, rate=5);
}
hist(replicates)
abline( v=5, col='red', lwd=4 )
abline(v=mean(replicates), col="black", lwd=4, lty=2)
```
```{r}
mean(replicates)-5

```

How does the bias change over increasing sample sizes? Does the bias go to zero?
How does the standard deviation change?

```{r}
ns <- seq(25,1000, by=50)
sd <- vector(mode="numeric")
bias <- vector(mode="numeric")

for(k in 1:length(ns)){
  M <- 1e4;
  replicates <- rep(NA, M);
  for (i in 1:M ) {
    replicates[i] <- run_exprate_expt(ns[k], rate=5);
  }
  sd[k] <- sd(replicates)
  bias[k] <- mean(replicates)-5
}
plot(x=ns,y=bias, type="l", ylim=c(0,bias[1]), main="Bias of 1/xbar as n increases")
abline(h=0)


plot(x=ns,y=sd, type="l", main="Standard error of 1/xbar as n increases")
lines(x=ns, y=sqrt(ns[1])*sd[1]/sqrt(ns), lty=2)
```


