---
title: "Homework 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
```

## Problem 1: The infamous mule kick data <small>10pts</small>

The file `mule_kicks.csv`, available for download (here)[https://kdlevin-uwstat.github.io/STAT340-Fall2021/hw/03/mule_kicks.csv], contains a simplified version of a very famous data set. The data consists of the number of soldiers killed by being kicked by mules or horses each year in a number of different companies in the Prussian army near the end of the 19th century.

This may seem at first to be a very silly thing to collect data about, but it is a very interesting thing to look at if you are interested in rare events. Deaths by horse kick were rare events that occurred independently of one another, and thus it is precisely the kind of process that we might expect to obey a Poisson distribution.

Download the data and read it into R by running

```{r}
download.file('https://kdlevin-uwstat.github.io/STAT340-Fall2021/hw/03/mule_kicks.csv', destfile='mule_kicks.csv')
mule_kicks = read.csv('mule_kicks.csv', header=TRUE)

head(mule_kicks)
```

`mule_kicks` contains a single column, called `deaths`.
Each entry is the number of soldiers killed in one corps of the Prussian army in one year.
There are 14 corps in the data set, studied over 20 years, for a total of 280 death counts.


### Part a: estimating the Poisson rate <small>3pts</small>

Assuming that the mule kicks data follows a Poisson distribution, produce a point estimate for the rate parameter $\lambda$.
There are no strictly right or wrong answers, here, though there are certainly better or worse ones.

```{r}
lambda_estimate <- mean(mule_kicks$deaths)
lambda_estimate
```


### Part b: constructing a CI <small>4pts</small>

Using everything you know (Monte Carlo, CLT, etc.), construct a confidence interval for the rate parameter $\lambda$.
Explain in reasonable detail what you are doing and why you are constructing the confidence interval in this way (a few sentences is fine!).

```{r}
# Number of observations
n <- length(mule_kicks$deaths)

# Standard error using lambda estimate
se <- sqrt(lambda_estimate / n)

# Critical value for 95% confidence level
z_value <- qnorm(0.975)  # 1.96 for a 95% CI

# Confidence interval
ci_lower <- lambda_estimate - z_value * se
ci_upper <- lambda_estimate + z_value * se

ci_lower
ci_upper
```


### Part c: assessing a model <small>3pts</small>

Here's a slightly more open-ended question.
We *assumed* that the data followed a Poisson distribution.
This may or may not be a reasonable assumption.
Use any and all tools that you know to assess (either with code or simply in words) how reasonable or unreasonable this assumption is.

Once again, there are no strictly right or wrong answers here.
Explain and defend your decisions and thought processes in a reasonable way and you will receive full credit.

```{r}
# Calculate mean and variance
sample_mean <- mean(mule_kicks$deaths)
sample_variance <- var(mule_kicks$deaths)

sample_mean
sample_variance

```

> A key property of the Poisson distribution is that the mean should equal the variance (lamda). Calculating these values for the deaths column, we find that they are quite similar.

```{r}
ggplot(data.frame(deaths), aes(x = deaths)) +
  geom_histogram(aes(y = ..density..), bins = 10, fill = 'skyblue', alpha = 0.7) +
  stat_function(fun = dpois, args = list(lambda = mean(deaths)), color = 'red', size = 1) +
  labs(title = "Histogram of Deaths with Poisson Fit",
       x = "Number of Deaths",
       y = "Density")
```

> When we graph the data, we can see that the histogram closely follows the red Poisson curve, which again supports the assumption.

```{r}
# Observed counts and expected frequencies
counts <- table(mule_kicks$deaths)
expected_counts <- dpois(as.numeric(names(counts)), lambda = sample_mean) * length(mule_kicks$deaths)

counts
expected_counts

# Chi-squared test
chisq_test <- chisq.test(counts, p = expected_counts, rescale.p = TRUE)
chisq_test

```

> With a Chi-squared statistic of 2.1498 and a p-value of 0.7082, we do not have sufficient evidence to reject the Poisson distribution as a reasonable model for this data.


## Problem 2: A Confidence interval for the median (8 points)

The following numbers were sampled from a population. You can assume they are independent and identically distributed (they are, in fact).

```{r}
mysteryData <- c(16.91,19.96,16.55,16.77,19.64,16.87,16.50,14.19,16.54,18.64,11.27,17.08) 
```

We want to estimate the population median with a 95\% confidence interval.

a. First assume that the population is normally distributed. Come up with estimates for the parameters of the normal distribution based on the data.

> In normal distribution the mean equals the median.

```{r}
mean_estimate <- mean(mysteryData)
sd_estimate <- sd(mysteryData)

mean_estimate
sd_estimate
```

b. Construct a 95\% confidence interval assuming the population is a normal distribution. Use the Monte Carlo Approach. Use 1000 Monte Carlo replicates.

```{r}
set.seed(123)
n_replicates <- 1000
sample_size <- length(mysteryData)

# Monte Carlo simulation
monte_carlo_medians <- replicate(n_replicates, {

  simulated_sample <- rnorm(sample_size, mean = mean_estimate, sd = sd_estimate)

  median(simulated_sample)
})

ci_lower <- quantile(monte_carlo_medians, 0.025)
ci_upper <- quantile(monte_carlo_medians, 0.975)

ci_lower
ci_upper
```

c. Now assume that the population is uniformly distributed. Come up with estimates for the min and max of the uniform distribution using the method of moments. This is based on $\bar{X}$, the sample mean and $S$, the sample standard deviation. The point estimates for a and b are given by  $$\bar{X} \pm \sqrt{3} S$$ 

```{r}
a_estimate <- mean_estimate - sqrt(3) * sd_estimate
b_estimate <- mean_estimate + sqrt(3) * sd_estimate

a_estimate
b_estimate
```

d. Construct a 95\% confidence interval assuming the population is a uniform distribution. Use 1000 Monte Carlo replicates.

```{r}
set.seed(123)
n_replicates <- 1000

monte_carlo_a <- numeric(n_replicates)
monte_carlo_b <- numeric(n_replicates)

for (i in 1:n_replicates) {

  simulated_sample <- runif(length(mysteryData), min = a_estimate, max = b_estimate)
  
  sample_mean <- mean(simulated_sample)
  sample_sd <- sd(simulated_sample)
  
  monte_carlo_a[i] <- sample_mean - sqrt(3) * sample_sd
  monte_carlo_b[i] <- sample_mean + sqrt(3) * sample_sd
}

ci_a_lower <- quantile(monte_carlo_a, 0.025)
ci_a_upper <- quantile(monte_carlo_a, 0.975)

ci_b_lower <- quantile(monte_carlo_b, 0.025)
ci_b_upper <- quantile(monte_carlo_b, 0.975)

ci_a_lower
ci_a_upper
ci_b_lower
ci_b_upper
```

## Problem 3: Closing the loop <small>6 pts</small>

In our discussion of the Universal Widgets of Madison company from lecture, we said that we were interested in two questions:

1. Estimating the probability $p$ that a widget is functional.
2. How many widgets should be in a batch to ensure that (with high probability) a batch ships with at least $5$ functional widgets in it?

We discussed question (1) at length in lecture.
What about question (2)?
Our client wants to know how many widgets should ship in each batch so as to ensure that the probability there are at least $5$ functional widgets in a batch is at least $0.99$.

Now, suppose that we have observed data and estimated $p$ to be $0.82$.

Use everything you know so far in this course to give a recommendation to the client.
Be sure to explain clearly what you are doing and why.
If there are any steps, assumptions, etc., that you are not 100% pleased with, feel free to point them out.

__Note:__ there are at least two "obvious" ways to solve this problem. One is based on using Monte Carlo (i.e., assume $p=0.82$ is the truth, and try generating batches of different sizes, etc.).
The other uses direct computation of probabilities, using basic facts about Binomial RVs.
Neither of these is necessarily better than the other, and you do not need to use both approaches to receive full credit.
Indeed, you are free to try doing something else entirely, if you wish.
Just explain clearly what you are doing and why!

> Approach with Monte Carlo:

```{r}
set.seed(123)

p <- 0.82

# Function to simulate a batch of widgets and check if at least 5 are functional
simulate_batch <- function(batch_size) {
  # Generate a batch of widgets, 1 represents functional, 0 non-functional
  batch <- rbinom(batch_size, 1, p)
  # Check if at least 5 widgets are functional (i.e., sum of 1's in the batch)
  return(sum(batch) >= 5)
}

# Function to estimate the minimum batch size for 99% probability
find_batch_size <- function(target_prob, max_size = 200) {
  for (batch_size in 5:max_size) {
    # Simulate 10000 batches and check the proportion with at least 5 functional widgets
    results <- replicate(10000, simulate_batch(batch_size))
    
    # Calculate probability of having at least 5 functional widgets in a batch
    prob_at_least_5 <- mean(results)
    
    cat("Batch size:", batch_size, "Probability of at least 5 functional widgets:", prob_at_least_5, "\n")
    
    if (prob_at_least_5 >= target_prob) {
      return(batch_size)
    }
  }
  return(NA) # if no size meets the probability
}

# Find the minimum batch size that ensures at least 5 functional widgets with 99% probability
min_batch_size_mc <- find_batch_size(0.99)
cat("Monte Carlo Approach - Minimum batch size:", min_batch_size_mc, "\n")

```

> Approach with Binomial Distribution:

```{r}
p <- 0.82 
required_prob <- 0.99 

# Function to calculate the minimum batch size
min_batch_size <- function(p, required_prob) {
  for (n in 1:1000) { 
    prob <- 1 - pbinom(4, n, p)  # Probability of having at least 5 functional widgets
    if (prob >= required_prob) {
      return(n)
    }
  }
  return(NULL)
}

# Calculate the minimum batch size
batch_size <- min_batch_size(p, required_prob)

cat("The minimum batch size to ensure at least 5 functional widgets with a probability of at least", required_prob, "is", batch_size, ".\n")
```

> Based on both the Monte Carlo and Binomial Distribution methods, the minimum batch size required to ensure that there are at least 5 functional widgets with a probability of at least 0.99 is 10.


## Problem 4: Estimating Coverage Rate (6 points)

Suppose we want to estimate the $p$ parameter from a Geometric distribution. By doing some research you would find that an unbiased estimator for $p$ is given as:

$$\hat{p} = \frac{1}{1+\frac{n}{n-1}\bar{X}}$$

a. Create a plot that shows the coverage rate of a 95\% confidence interval for samples of size 12 for values of p: .10, .20, ..., .80, .90. 

* To estimate the coverage rate, use 500 Monte Carlo replicates
* To construct each MC interval you should use 500 MC replicates

If this is taking too long to run, lower both of these to 200 replicates.

```{r}
set.seed(1)
p <- seq(.1,.9,.1)
coverage <- rep(0, length(p))
results <- data.frame(p, coverage)
NMC1 <- 500
NMC2 <- 500

for(j in 1:nrow(results)){
  # pull out the value p from the results data frame
  p <- results$p[j]
  # the 'contains' vector will be a collection of TRUE and FALSE, based
  # on whether each CI contains the value p
  contains <- rep(FALSE, NMC1)
  for(i in 1:NMC1){ #Construct NMC1 confidence intervals
    #generate some data initially from the geometric distribution

    #calculate the unbiased estimate of p
    
    #Now you want to use your p.hat to generate NMC2 sets of data
    #Each time you want to calcualte an unbiased phat, and store it.
    
    #Your Monte Carlo Confidence interval should be based on
    #appropriate sample quantiles
    
    #Last, assign 'contains[i]' to be TRUE or FALSE based on 
    #whether p is >= than the lower bound AND <= the 
    #upper bound. 
  }
  results[j, "coverage"] <- mean(contains)
}
plot(x=results$p, y=results$coverage, ylim=c(min(coverage),1), type="l")
abline(h=.95, lty=2)
```

b. What do you observe? Does the coverage of MC intervals tend to be affected by the value of $p$? In what way?
