---
title: "Bootstrapping Examples"
author: "Brian Powers"
date: "2024-02-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bootstrapping the parameter p for a Geometric population
Suppose we have a geometric random variable (i.e population) with some unknown p
```{r}
p <- runif(1)
```
Modify this code to a non-parametric method.

Please note that this is not a very good way to use bootstrapping - we will learn a better way later this semester. This is only to highlight how we can still estimate a confidence interval even if we do not know what population model to use.
```{r}
# Bootstrapping
NMC2 <- 1000
correct.count <- 0
for(j in 1:NMC2){
  my.sample <- rgeom(n=50, prob=p)
  NMC <- 1000
  p.hats <- vector("numeric")
  for(i in 1:NMC){
    simulated.sample <- sample(x=my.sample, size=length(my.sample), replace=TRUE)
    p.hats[i] <- 1 / (mean(simulated.sample) + 1)
  }
  #Let's do a 90% confidence interval for the population parameter p
  bounds <- quantile(p.hats, probs = c(.05, .95))
  correct.count <- correct.count + as.numeric(bounds[1] <= p & bounds[2] >=p)
}
correct.count / NMC2
```

Bootstrapping a normal population
```{r}
NMC2 <- 1000
correct.count <- 0
mu <- runif(1, 0,100)
sigma <- runif(1, 5,10)
for(j in 1:NMC2){
  my.sample <- rnorm(n=25, mean=mu, sd=sigma)
  NMC <- 1000
  x.bars <- vector("numeric")
  for(i in 1:NMC){
    simulated.sample <- sample(x=my.sample, size=length(my.sample), replace=TRUE)
    x.bars[i] <- mean(simulated.sample)
  }
  #Let's do a 90% confidence interval for the population parameter p
  bounds <- quantile(x.bars, probs = c(.05, .95))
  correct.count <- correct.count + as.numeric(bounds[1] <= mu & bounds[2] >=mu)
}
correct.count / NMC2
```

