---
title: "R11 Model Selection"
author: "Brian Powers"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Selection in mtcars

In this code we will compare different multiple regression models to predict miles per gallon from the various predictor variables in the mtcars dataset.

We implement a forward stepwise selection algorithm, and a backwards stepwise selection algorithm. We also have a  function that automates K fold validation and gives the formula resulting in the lowest MSE


### K fold validation on the models produced

```{r, warning=FALSE}
kfoldCV <- function(K, formulas, dataset){
  m <- length(formulas)
  
  #idx is a shuffled vector of row numbers
  idx <- sample(1:nrow(dataset))
  #folds partitions the row indices
  folds <- split(idx, as.factor(1:K))

  #an empty data frame to store the results of each validation
  results <- data.frame(fold = rep(1:K, rep(m,K)),
                        model = rep(1:m, K),
                        error = 0)    
  for(k in 1:K){
    #split the data into training and testing sets
    training <- dataset[-folds[[k]],]
    testing <- dataset[folds[[k]],]
    #go through each model and estimate MSE
    for(f in 1:m){
      #fit the model to the training data
      fit <- lm(formula = formulas[[f]], data=training)
      #calculate the average squared error on the testing data
      results[results$fold == k & results$model == f, "error"] <- mean((predict(fit, newdata=testing)-testing$mpg)^2)
    }
  }
  #aggregate over each model, averaging the error
  aggregated <- aggregate(error~model, data=results, FUN="mean")
  #produces a simple line & dot plot
  plot(error ~ model, type="b", data=aggregated, ylab="MSE")
#  lines(error ~ model, data=aggregated)
  print(which(aggregated$error == min(aggregated$error)))
  print(formulas[[which(aggregated$error == min(aggregated$error))]])
  return(aggregated)
}
```


### Model Statistics

```{r}
createModelStats <- function(formulas, dataset){
  n <- nrow(dataset)
  model.stats <- data.frame(model <- as.character(formulas))
  for(m in 1:length(formulas)){
    modelsum <- summary(lm(formulas[[m]], data=dataset))
    #I am using k = p+1, the number of predictors + the intercept
    model.stats$k[m] <- nrow(modelsum$coefficients)
    model.stats$Rsq[m] <- modelsum$r.squared
    model.stats$Rsq.adj[m] <- modelsum$adj.r.squared
    L <- logLik(lm(formulas[[m]], data=mtcars))[1]
    model.stats$AIC[m] <- -2*L + 2*model.stats$k[m]
    model.stats$BIC[m] <- -2*L + model.stats$k[m] * log(n)
  }
  return(model.stats)  
}
```



### Forward Stepwise Selection

```{r}
predictors = c("1",names(mtcars)[2:11])
used = 1
M <- list(10)


M0 <- lm(reformulate(predictors[used], "mpg"), data=mtcars)

RSS <- anova(M0)["Residuals","Sum Sq"]

formulas <- list()

for(model in 1:10){
  RSS.best <- RSS
  for(try in predictors[-used]){
    M <- lm(reformulate(c(predictors[used],try), "mpg"), data=mtcars)
    RSS.new <- anova(M)["Residuals","Sum Sq"]
    if(RSS.new <= RSS.best){
      new.pred <- try
      RSS.best <- RSS.new
    }
  }
  formulas[[model]] <- reformulate(c(predictors[used],new.pred), "mpg")
  M[[model]] <- lm(formulas[[model]], data=mtcars) 
  RSS <- anova(M[[model]])["Residuals","Sum Sq"]
  print(paste("adding", new.pred, "; RSS = ", RSS))
  used <- c(used, which(predictors==new.pred))
}

kfoldCV(4, formulas, mtcars)

createModelStats(formulas, mtcars)
```





### Backwards Stepwise

```{r}
M <- list(10)

predictors = c("1",names(mtcars)[2:11])
used = 2:11


M10 <- lm(reformulate(predictors[c(1,used)], "mpg"), data=mtcars)
RSS <- anova(M10)["Residuals","Sum Sq"]
formulas <- list()
formulas[[10]] <- reformulate(predictors[used], "mpg")

print(paste("Full Model RSS: ", RSS))
for(model in 9:1){
  RSS.best <- var(mtcars$mpg)*(nrow(mtcars)-1)
  for(try in used){
    M <- lm(reformulate(predictors[used[-which(used==try)]], "mpg"), data=mtcars)
    RSS.new <- anova(M)["Residuals","Sum Sq"]
    if(RSS.new <= RSS.best){
      new.pred <- try
      RSS.best <- RSS.new
    }
  }
  formulas[[model]] <- reformulate(predictors[used[-which(used==try)]], "mpg")
  M[[model]] <- lm(formulas[[model]], data=mtcars) 
  RSS <- anova(M[[model]])["Residuals","Sum Sq"]
  print(paste("removing", predictors[new.pred], "; RSS = ", RSS))
  used <- used[-which(used==new.pred)]
}

kfoldCV(8, formulas, mtcars)
createModelStats(formulas, mtcars)

```


