---
title: "Homework 2"
output: 
  html_document:
    self_contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
```


## Problem 1 <small>(2pts each, 8pts total)</small>

Here are a few probability exercises to get you warmed up.

a. Consider $X\sim \text{Normal}(5, 2^2)$. Find $P(4 < X < 7)$
```{r}

pnorm(7, 5, 2) - pnorm(4, 5, 2)

```

b. Suppose you have an uneven 6-sided die where the even numbers are equally likely to occur and the odd numbers are also equally likely. The expected value of the die is $3.3333$ (i.e. $\frac{10}{3})$. What is the probability of getting a 6?
   (Hint: Let $p$ represent the probability of getting an even number, and represent the probability of the other numbers algebraically. Then derive an expression for the expected value, let it equal $\frac{10}{3}$, and solve. You can go back and check your work by plugging the result back into your original expressions and verify it works with R.)
```{r}
E = 10/3

# E = sum( p *   )


```

c. Give **two examples** of pairs of events that are mutually exclusive and explain why for each.

> One example of mutually exclusive events is: rolling the number 5 with a fair 6 sided die, and the number you roll being even. Explanation: 5 is not an even number.

> Another example is: get a heads after flipping a coin, and getting tails after flipping a coin. Explanation: when flipping a coin, there is only one outcome between heads or tails.

d. Give **two examples** of pairs of events that are independent and explain why for each.

> Rolling the number 5 with a fair 6 sided die, and drawing the card 5 from a normal deck of cards.

> Get a heads when flipping a coin, and probability of raining that day.

## Problem 2 <small>(2pts each, 14pts total)</small>

For each of the following scenarios, decide what random variable may be appropriate to use as a model, and explain your reasoning as well as any assumptions you make.

a. The number of patients in an experimental drug trial that respond to treatment. 

> Binomial Distribution

b. The number of red cars you observe driving through an intersection between 10am and 11am. 

> Poisson Distribution

c. What the second hand reads on the clock when you wake up from a long nap (approximately). 

> Uniform Distribution

d. How many people you need to swipe right on Tinder before you get a match. 

> Geometric Distribution

e. The length of time between mosquito bites a camper experiences while on a hike. 

> Exponential Distribution

f. Whether the Eagles win the Superbowl this year. 

> Binomial (Bernoulli) Distribution

g. The GPA of a randomly selected UW-Madison graduate.

> Normal Distribution


## Problem 3 <small>(2pts each, 4pts total)</small>

a. Consider the CDF given below for random variable $Y$ 
```{r}
plot(x=0:6, y=c(0,.3,.5, .6, .75,.8,1), type="s", ylab="F(y)", xlab="y", main="cdf of Y")
abline(h=seq(.1,.9,.1), lty=2, col="gray")
```
Express the pmf of $Y$ in the following R code (in other words, fill in the missing probabilities)

```{r}
y<- 1:6 ## 1 do 6 ima samo 6 elementi, a sose nula ima 7 elementi
py <- c(0,0.3,0.8,1.40,2.15,2.95) ## edit this line
barplot(height=py, names=y, main="pmf of Y")
```



b. Calculate $E[Y]$ for the variable $Y$ in part a.

> tuka mi treba probability za da presmetam?

## Problem 4 <small>(2pts each, 4pts total)</small>

Below you will find the plot of a probability density function for a random variable $X$.

```{r, echo=FALSE}
plot(x=c(0,1,2,3), y=c(0,.5,.5,0), type="l", ylab="density", xlab="x", main="pdf of X")
```

Using just geometry answer a and b:

a. What is $Pr[X < 1]$?
> 1 * 0.5 / 2 = 0.25

b. What is $Pr[1.5 < X < 2.5]$
> (2-1.5) * 0.5 + (2.5-2) * 0.25 = 0.25 + 0.125 = 0.375



