---
title: "Homework 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
```

## Problem \#1: Testing coin flips <small>(6 pts)</small>

In the six sequences below, only one of them is actually **randomly generated from independent flips of a fair coin**. Use a combination of everything you know (common sense, Monte Carlo, hypothesis testing, etc.) to identify which is actually random and explain your reasoning.

(For full points, conduct a formal test and report a $p$-value for each sequence. You may use a combination of multiple tests to arrive at your answer. If you cannot compute a p-value for each sequence, you can still earn a significant amount of partial credit by carefully explaining your reasoning and response as best as you can.)

My advice is **be creative** with the test statistics you come up with to eliminate each sequence! Think of some way of summarizing a sequence of flips that might be useful for comparing against a simulated sequence of random flips. After you come up with an idea for a statistic, remember to run it on many MC generated completely random flips to produce a distribution under the null, which you can then compare with your data to get a p-value. Also, be careful of now you define "more extreme" than the data.

(2 bonus points available if you can find a single test that is powerful enough to reject all the fake sequences together in one step. Yes, at least one such possible test exists.)

```{r}
flips1 = "HTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHT"

flips2 = "HHHTHTTTHHTHHTHHHTTTTHTHTHHTTHTHHHTHHTHTTTHTHHHTHTTTHTHTHHTHTHTTHTHHTHTHTTTHTHHHTHTHTTHTHTHHTHTHTHHHTHTTTHTHHTHTHTHHTTTHTHHTHHTTTTHTHTHHHTHTTHTHHTHTHTTHTHHTHTHHHTHHHTHTTTHTTHTTTHTHHHTHTHTTHTHHTHHTHTTT"

flips3 = "HHTHTHTTTHTHHHTHHTTTHTHHTHTTTHTHTHHTHTHTTHTHHHHHHTTTHTHTHHTHTTTHTHHTHTHTTTHTHHHTTHTTTHTHTHHHHTHTTHHTTTTTHTHHHTHTHTTTTTHHHTHHTHHTHHHTTTTHTHTHHHTHHTTTTTHTHHHTHTHTHTTTHTHHHTHTHTHTTHTHHTHTHTHTTTTHTHHHTHTH"

flips4 = "HTHHHHHHHTHTTHHTTHHHTHTHTTTHHTHHHTHHTTHTTTTTTTTTHTHHTTTTTHTHTHTHHTTHTTHTTTTTHHHTHTTTHTHTHHHTHTTTTHTHTHHTTHTHTTHHTHTHHHHTHTTHHTTHTTHTTHTHHHHHHTTTTTTHHHTTHTHHHHTTTHTTHHHTTHTHHTTTHHTHHTTTHTHHTHHHTHHTTHHH"

flips5 = "HHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTT"

flips6 = "TTHTTTHTTTTTTTHTHTHTHTTHTTHTHHTHHTTTHHTHTTTHTHHTHHHTHTTHHTHHTTHTHTTTTHTHTTTHHTTTTTTTTHTHHTTHTTTTTTHTHTHTHTTTHTTHHTTHTTTHHTTTHTTHTTTTHTTTTHHTTTHTHTHHHTTTTTTHTHHTTTTTTTTTTTTHHHTTTHHHTTTHTTTHTHTTHTTTTTHT"

# you can use the function below to split the above sequences in vectors of flips
split = function(str) strsplit(str, split="")[[1]]
split(flips1)

# Function to calculate autocorrelation for a given lag
autocorrelation <- function(sequence, lag) {
  n <- length(sequence)
  sequence_lagged <- sequence[(lag + 1):n]
  sequence_original <- sequence[1:(n - lag)]
  
  # Compute correlation between original and lagged sequence
  return(cor(sequence_original, sequence_lagged))
}

# Function to calculate the maximum streak length
max_streak_length <- function(sequence) {
  max_streak <- 0
  current_streak <- 1
  
  for (i in 2:length(sequence)) {
    if (sequence[i] == sequence[i - 1]) {
      current_streak <- current_streak + 1
    } else {
      max_streak <- max(max_streak, current_streak)
      current_streak <- 1
    }
  }
  max_streak <- max(max_streak, current_streak) # Handle case when last streak is the longest
  return(max_streak)
}

# Convert coin flips sequences to 0s and 1s (H = 1, T = 0)
convert_to_numeric <- function(sequence) {
  return(as.numeric(sequence == 'H'))
}

# Example sequences
flips1 <- unlist(strsplit("HTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHTHT", ""))
flips2 <- unlist(strsplit("HHHTHTTTHHTHHTHHHTTTTHTHTHHTTHTHHHTHHTHTTTHTHHHTHTTTHTHTHHTHTHTTHTHHTHTHTTTHTHHHTHTHTTHTHTHHTHTHTHHHTHTTTHTHHTHTHTHHTTTHTHHTHHTTTTHTHTHHHTHTTHTHHTHTHTTHTHHTHTHHHTHHHTHTTTHTTHTTTHTHHHTHTHTTHTHHTHHTHTTT", ""))
flips3 <- unlist(strsplit("HHTHTHTTTHTHHHTHHTTTHTHHTHTTTHTHTHHTHTHTTHTHHHHHHTTTHTHTHHTHTTTHTHHTHTHTTTHTHHHTTHTTTHTHTHHHHTHTTHHTTTTTHTHHHTHTHTTTTTHHHTHHTHHTHHHTTTTHTHTHHHTHHTTTTTHTHHHTHTHTHTTTHTHHHTHTHTHTTHTHHTHTHTHTTTTHTHHHTHTH", ""))
flips4<- unlist(strsplit("HTHHHHHHHTHTTHHTTHHHTHTHTTTHHTHHHTHHTTHTTTTTTTTTHTHHTTTTTHTHTHTHHTTHTTHTTTTTHHHTHTTTHTHTHHHTHTTTTHTHTHHTTHTHTTHHTHTHHHHTHTTHHTTHTTHTTHTHHHHHHTTTTTTHHHTTHTHHHHTTTHTTHHHTTHTHHTTTHHTHHTTTHTHHTHHHTHHTTHHH", ""))
flips5<- unlist(strsplit("HHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTTHHHHHHHHHHTTTTTTTTTT", ""))
flips6<- unlist(strsplit("TTHTTTHTTTTTTTHTHTHTHTTHTTHTHHTHHTTTHHTHTTTHTHHTHHHTHTTHHTHHTTHTHTTTTHTHTTTHHTTTTTTTTHTHHTTHTTTTTTHTHTHTHTTTHTTHHTTHTTTHHTTTHTTHTTTTHTTTTHHTTTHTHTHHHTTTTTTHTHHTTTTTTTTTTTTHHHTTTHHHTTTHTTTHTHTTHTTTTTHT", ""))

# Convert to numeric
flips1_num <- convert_to_numeric(flips1)
flips2_num <- convert_to_numeric(flips2)
flips3_num <- convert_to_numeric(flips3)
flips4_num <- convert_to_numeric(flips4)
flips5_num <- convert_to_numeric(flips5)
flips6_num <- convert_to_numeric(flips6)


# Calculate autocorrelation for Lag 1 and Lag 2
autocorr_flips1_lag1 <- autocorrelation(flips1_num, 1)
autocorr_flips1_lag2 <- autocorrelation(flips1_num, 2)
autocorr_flips2_lag1 <- autocorrelation(flips2_num, 1)
autocorr_flips2_lag2 <- autocorrelation(flips2_num, 2)

# Calculate max streak length
max_streak_flips1 <- max_streak_length(flips1_num)
max_streak_flips2 <- max_streak_length(flips2_num)
max_streak_flips3 <- max_streak_length(flips3_num)
max_streak_flips4 <- max_streak_length(flips4_num)
max_streak_flips5 <- max_streak_length(flips5_num)
max_streak_flips6 <- max_streak_length(flips6_num)

# Results
cat("Flips1 Autocorrelation (Lag 1):", autocorr_flips1_lag1, "\n")
cat("Flips1 Autocorrelation (Lag 2):", autocorr_flips1_lag2, "\n")
cat("Flips2 Autocorrelation (Lag 1):", autocorr_flips2_lag1, "\n")
cat("Flips2 Autocorrelation (Lag 2):", autocorr_flips2_lag2, "\n")
cat("Flips1 Max Streak Length:", max_streak_flips1, "\n")
cat("Flips2 Max Streak Length:", max_streak_flips2, "\n")
cat("Flips3 Max Streak Length:", max_streak_flips3, "\n")
cat("Flips4 Max Streak Length:", max_streak_flips4, "\n")
cat("Flips5 Max Streak Length:", max_streak_flips5, "\n")
cat("Flips6 Max Streak Length:", max_streak_flips6, "\n")

# Function to simulate coin flips and find max streak length
simulate_max_streak <- function(n_flips) {
  sequence <- sample(c(0, 1), n_flips, replace = TRUE)  # Random sequence (0 = T, 1 = H)
  max_streak_length(sequence)
}
simulate_max_streak(200)
# Simulate max streak lengths for many random sequences

n_simulations <- 10000
n_flips <- 200
simulated_streaks <- replicate(n_simulations, simulate_max_streak(n_flips))

# Compare observed max streaks with simulated distribution
observed_streaks <- c(1, 4, 6, 9, 10, 12)

# Calculate the proportion of simulated streaks that are longer than or equal to each observed streak
p_values <- sapply(observed_streaks, function(obs) mean(simulated_streaks >= obs))

# Results
cat("P-values for each observed streak length:", p_values, "\n")






```

> Here 

## Problem \#2: Finding the Trick Coin <small>(6 pts; 2pts each)</small>

I have two coins in my pocket - a trick coin with two heads and a fair coin with one head and one tail(s?). We'll play a game. I will grab one coin at random, and flip it $N$ times. After that you will have to decide if it is the fair coin or the trick coin. The null hypothesis is that it is the fair coin. 

**Decision Rule 1**: If after $N$ flips there are no tails, then you decide it is the trick coin. If there is at least 1 tail then you know it is the fair coin. 

a. Using "Decision Rule 1", what is the lowest number of flips $N$ would you need in order to have a significance level less than 5% for this test?
b. Using $N$ from part a, what is the power of the test?
c. Suppose $N=4$ is decided. How can you modify the decision process to have a significance level of exactly 5%? Does this change the power of the test?
d. Extra Credit (2 points): Suppose if you guess correct you win \$100 (and if you're wrong you get nothing), but each flip of the coin costs \$10. What strategy would you use to maximize your expected profit from this game?


## Problem \#3: Testing the maximum of a uniform distribution <small>(8 pts; 2 pts each)</small>

We sample $X_1, X_x,\ldots,X_n \overset{\text{iid}}\sim\text{Uniform}(0,m)$ where $m$ is an unknown maximum. Sleazy Jim tells you that $m=1$ but you're not so sure. The 50 values sampled are in the following data file:

```{r}
X <- read.csv("uniform_sample.csv")$x
```

a. Write out in formal notation the null and alternative hypotheses. 
b. Come up with a test statistic and measure your sampled data. Is this a one-sided test or two-sided test?
c. Simulate a distribution for the test statistic under the null hypothesis of size at least 1000. Display a histogram of your test statistic distribution.
d. Calculate the $p$-value for this data and make a conclusion.

## Problem \#4: Blurtle <small>(10 pts; 2 pt each)</small>

Have you been playing the hot new game Blurtle? It's a (fictional) word game you can play daily - you have to guess a 5 letter word and you only have 6 attempts. I've been playing for the past 100 days and I've been tracking my number of guesses. I'm trying to figure out whether I have been getting better or not.

The file `blurtle.csv` contains 100 rows of data, giving the number of tries to guess the word. If it took 7 guesses that actually means I failed (you don't actually get a 7th guess). 

```{r}
blurtle <- read.csv("blurtle.csv")
```

Your task is to perform a permutation test on the data to determine if there is statistical evidence of a true improvement trend.

a. State the null and alternative hypotheses
b. Determine a test statistic that identifies a trend in the number of tries. There are many good ones you could use - be creative.
c. Decide whether the test will be a one or two-tailed test
d. Simulate a distribution of test statistics under the null hypothesis
e. Calculate the test statistic on the observed data, calculate the $p$-value and state your conclusions.


