---
title: "FA24 STAT340 Midterm <small>take-home portion</small>"
output: html_document
---

<style>h2{margin-top:5rem}h3,h4:not([class]){margin-top:4rem}
.main-container{padding:2rem}</style>
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,warning=F,message=F,
                      fig.width=5,fig.height=4,fig.align="center")
```


## Q1 - Testing

Infinity Cable (fictional) manufactures fiber optic cable for industrial purposes. They have three manufacturing facilities in Alderwood, Brentwick and Coppergrove. A recent quality control audit sampled various 1 kilometer lengths of cable from the three facilities and counted the number of flaws in the cables:

```{r}
cables = data.frame(
  site = rep(c("Alderwood","Brentwick","Coppergrove"), times=c(53,48,51)),
  flaws = c(6,4,4,1,2,1,4,2,4,3,4,8,2,4,6,2,4,1,2,2,0,2,5,2,3,3,3,2,5,4,4,
            1,4,3,5,4,4,3,3,4,0,3,4,4,3,5,3,2,1,1,2,3,4,2,5,1,2,2,3,1,2,3,
            0,4,2,4,2,2,2,4,4,2,3,5,2,3,2,2,3,1,3,1,1,1,1,0,3,4,4,4,2,2,4,
            3,3,2,1,7,3,1,1,3,6,3,7,4,2,3,1,0,4,1,3,4,8,3,3,1,4,3,3,2,2,3,
            3,1,0,4,6,3,3,3,7,3,4,3,2,2,4,3,1,4,1,5,3,3,2,3,3,1,3,1)
)
```

Your job is to perform a comparison using the testing techniques we've covered in this course to see if there is evidence that the facilities actually produce cable of different quality. Do not use techniques we have not covered in this course.

a. (2pts) A colleague suggests that we should focus our attention on a comparison of Alderwood and Brentwick (ignoring Coppergrove), since they seem to have the largest gap in average number of flaws. Why specifically would this not be a good approach?

b. Now you will perform the test. You should

   i. (1pt) State your null and alternative hypotheses, 
   ii. (1pt) decide on the test statistic to use, 
   iii. (1pt) state whether this is a one or two-sided test, and explain why, 
   iv. (2pts) produce a simulated distribution for your test statistic under the null hypothesis 
   v. (2pts) calculate the $p$-value
   vi. (1pt) make your conclusion

*Hint: your test statistic should somehow compare all 3 groups. There are many possible statistics you could use. If you are stuck consider a statistic based on the site means or the site variances.*

> ***PLEASE WRITE YOUR WORK BELOW:***


```{r}
# insert code here

```

*write discussion here*







## Q2 - Estimation

Let $X_i$ be i.i.d random variables following continuous Uniform($-a,a$) for some unknown positive parameter $a$. Suppose you draw the following sample of $n=15$ observations:

```{r}
data = c(-1.97, -1.07, 0.61, 3.43, -2.51, 3.35, 3.74, 1.35,
         1.08, -3.68, -2.47, -2.72, 1.57, -0.97, 2.27)
```

a. (3pts) Using what you know about estimation, find a reasonable point estimator $\hat a$. Explain why you think this gives a reasonable estimate of the parameter $a$.
   - Note: you do NOT need to find the BEST estimator to earn full points on this part (or later parts), just one that can be argued to be reasonable.
b. (2pts) Apply your estimator to the given sample data. What is your point estimate for $a$?
c. (3pts) Using the MC method shown in class, compute **and interpret** a 95\% confidence interval for $a$. Also, *briefly* (1-2 sentences) explain how well you think this interval would actually perform.
d. (2pts) Just for this part d, suppose the true value of $a$ is actually 3.9. Use MC to approximately estimate the bias of your estimator.

> ***PLEASE WRITE YOUR WORK BELOW:***


```{r}
# insert code here

```

*write discussion here*







## Q3 - Regression

In R, the built-in dataset `USArrests` contains 1975 estimates of rates (per 100,000 people) of various types of crimes in each US state (source: the World Almanac). Consider the `Assault` and `Murder` columns as X and Y variables:


```{r}
x = USArrests$Assault
y = USArrests$Murder
# plot(x,y)
```

a. (2pts) Does this look like a reasonable dataset to fit a linear model to? Explain why or why not.
b. (2pts) Regardless of your answer to part a, fit a linear model to this data. Show the summary print out with the full table of coefficients.
c. (2pts) Does this result show a significant relation between x,y? Answer why or why not and give a p-value.
d. (2pts) Give a 95% confidence interval for the true slope.
e. (2pts) Make the standard diagnostic plot(s) for the residuals. Does this fit look reasonable? Explain why or why not.

> ***PLEASE WRITE YOUR WORK BELOW:***


```{r}
library(ggplot2)
# insert code here
# Load the dataset
data("USArrests")
x = USArrests$Assault
y = USArrests$Murder
plot(x,y)

# Fit the linear model   Murder=β0+β1Assault
model <- lm(Murder ~ Assault, data = USArrests)

# Print the summary
summary(model)

# Calculate the 95% confidence interval for the slope
confint(model, level = 0.95)

#plot the line
pp <- ggplot( model,
              aes(x=USArrests$Assault,y=USArrests$Murder));
pp <- pp +geom_point() + geom_smooth(method="lm",
                                     formula="y~x",
                                     se=FALSE);
pp <- pp + labs( x="Assult)",
                 y="Murder",
                 title="Murder and Assult" )
pp

#residuals
resids <- residuals(model)
hist(resids)

#Homoscedasticity
plot(model, which=1)

plot(model, which=2)

#all standard diagnostic
par(mfrow = c(2, 2))
plot(model)



```

a. It looks resonable to fit a linea r model with positive koficient specially for smaller values of variable x=Assult. For bigger values dispersion will be greater.

b. (2pts) Regardless of your answer to part a, fit a linear model to this data. Show the summary print out with the full table of coefficients.

c. (2pts) Does this result show a significant relation between x,y? Answer why or why not and give a p-value.

The p value for intercep 0.464 is quite high, indicating that the intercept is not statistically significant. This means that the intercept (the expected value of Murder when Assault is zero) is not significantly diferentf from zero.
The p-value for the slope 2.6e-12 is extremely low, indicating that the Assault variable is highly statistically significant. This means there is a very strong evidence to suggest a relationship between Assault rates and Murder rates.

d. (2pts) Give a 95% confidence interval for the true slope.

95% Confidance interval for the slope  0.03284621 0.05097104

e. (2pts) Make the standard diagnostic plot(s) for the residuals. Does this fit look reasonable? Explain why or why not.
1. Residuals vs Fitted: From the linear plot checking of the linearity assumption shows that residuals are rather randomly scattered around the fitted line. 
2. Histogram of residuals with some longer right tail
3. Homosedacity  Rather linear around horizontal line y=0. Slightly higher variance for the bigger values tend to negative values 
4. Normal Q-Q:  The points rather follow the diagonal line especially for the central part. Small deviations showing non-normality for the right tail of distribution.


