---
title: "Homework 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T,eval=T,message=F,warning=F,fig.align='center')
library(tidyverse)
```


## Problem 1. A Data Scientist Referees Ping Pong <small>(8pts total)</small>

The game is Ping Pong. Players grab their paddles and hit the ping pong ball back and forth scoring points one at a time. The game continues until one player reaches 21 points, and at that point the game ends unless the point difference is less than 2. If it is less than 2 the game continues until one player wins by 2. 

Suppose Athena and Bacchus play and Bacchus wins 21 to 15. Bacchus is super excited but Athena says that they should have a rematch, because she's sure that Bacchus is not better than her, it was just a fluke. 

Time for a Data Scientist to settle this dispute. We must consider two hypotheses. The null hypothesis is that they are equally skilled - thus for each point scored the probability it goes to the ultimate winner is $0.50$ (this is basically what Athena is claiming). The alternative is that Bacchus truly is more skilled, and the probability is greater than $0.50$ (the winner actually has more skill, and doesn't win purely by chance).

Create a Monte Carlo simulation of a game. Use the point difference at game end as the test statistic. 

a. (4 points) Create a function called `playPingPong` which simulates a single game of Ping Pong with equally skilled players. Remember the logic : points are given to players A or B with equal chance, and that continues until (1) the max score >= 21 and (2) the difference between scores >=2. Have the function return the point difference.

```{r}
# Function to simulate a single Ping Pong game
playPingPong <- function(scoreTrigger = 21, winMargin=2, probPlayerA=0.5){
  #initialize a score vector to be 0,0
  score_A <- 0
  score_B <- 0
  #repeat while the winning condition has not been reached
  #i.e. repeat while max score < 21 or point difference < 2
   while (TRUE) {
    # Randomly assign a point to player A or B with equal chance
    if (runif(1) < probPlayerA) {
    #Assign a point to player 1 or 2 randomly
      score_A <- score_A + 1
    } else {
      score_B <- score_B + 1
    }
    # Check if the game conditions are met
    if ((score_A >= scoreTrigger || score_B >= scoreTrigger) && abs(score_A -  score_B) >=        winMargin) {
      # Return the point difference when the game ends
      return(abs(score_A - score_B))
    } 
  }
}
# Example of how to run the simulation
point_difference <- playPingPong()
cat("The point difference is:", point_difference, "\n")

```

b. (2 points) Perform the Monte Carlo test; simulate 1000 (or more) games with equally skilled players. Look at the distribution of point differences, and compare the observed point difference to this distribution. What is the *p*-value of the observed point difference.

```{r}
#TODO: Enter Code Here
# Monte Carlo simulation of 1000 games
set.seed(123)  # Set seed for reproducibility
n_simulations <- 1000
simulated_differences <- replicate(n_simulations, playPingPong())

# Plot distribution of simulated point differences
hist(simulated_differences, breaks = 30, main = "Distribution of Point Differences in 1000 Simulated Games",
     xlab = "Point Difference", col = "lightblue", border = "black")

# Define an observed point difference (e.g., from a real game or previous simulation   10?)
observed_point_difference <- 10

# Calculate the p-value: proportion of simulated games with a point difference >= observed
p_value <- mean(simulated_differences >= observed_point_difference)
cat("The p-value is:", p_value, "\n")
mean(simulated_differences)
```


c. (2 points) How do you conclude? Is this one game sufficient evidence that Bacchus is the superior Ping Pong Player?

The game does not provide enough statistical evidence to say that player A is superior to player B. The observed point difference of 10 could reasonably occur due to random chance if both players are equally skilled, as suggested by the large p-


## Problem 2: Quality or Quantity? <small>(6pts total)</small> 

Marcio Ranchello (fictional) is a prolific architect who has won many accolades. For example, in the ranking of "Best 10 designs of 2023", 4 of the 10 designs are from Marcio Ranchello. The authors of the ranking report suggest that this is evidence of his greatness. However, you notice that among the 150 buildings considered in the rankings, 30 of them were designed by Marcio. Indeed, Marcio leads a big architecture firm that has been extremely active in designing new buildings.

What do you think? Is the ranking evidence of the quality of his work, or a consequence of the quantity of his designs?

Take the null hypothesis to be that any of the 150 considered buildings could be included in the top 10 with equal likelihood. How likely under this model would we see 4 (or more) of Ranchello's buildings in the top 10? What do you conclude? 

Proceed by treating this as a formal hypothesis test. Define the null and alternative hypotheses, define your test statistic, produce a distribution of simulated test statistics from the null model and finish by calculating a p-value and providing your own interpretation.

Solution
This problem can be modeled using the hypergeometric distribution, which is used for sampling without replacement from a finite population. We are interested in the probability of selecting 4 or more buildings designed by the same person when randomly picking 10 buildings from the total set of 150 buildings, 30 of which belong to the same person.

The hypergeometric probability mass function is given by:
P(X=k)=C(30,k)*C(120,10-k)/(150,10)

where 
- X is the number of buildings built by the same person in the top 10.
- C(30,k) is the number of ways to choose k buildings from the 30 build by the person 
- C(120,10-k) is the number of ways to choose the remaining buildings from the 120 other buildings.
- C(150,10) is the total number of ways to choose 10 buildings from the full set of 150.
We want to calculate P(X>=4)

H₀ (Null): The presence of 4 or more of the person's buildings in the top 10 is due to chance.
H₁ (Alternative): The presence of 4 or more of the person's buildings in the top 10 indicates higher quality, not due to chance.


```{r}
# Set up parameters
total_buildings <- 150  # Total number of buildings
person_buildings <- 30  # Number of buildings by the person
top10_buildings <- 10   # Number of top 10 buildings
success_in_top10 <- 4   # At least 4 buildings by the person

# Calculate the probability of getting 4 or more of the person's buildings in the top 10
p_value <- 1 - phyper(success_in_top10 - 1, person_buildings, total_buildings - person_buildings, top10_buildings)

cat("The p-value is:", p_value, "\n")
```
The p-value suggests the ranking could be a result of chance, implying that submitting more designs increases the likelihood of some of them making it into the top 10.
We do not have enough evidence to reject H0.



```{r}
# Set up parameters
#total_buildings <- 150      # Total number of buildings
#person_buildings <- 30      # Buildings by the person
#top10_buildings <- 10       # Number of buildings selected for top 10
success_threshold <- 4      # We want to see at least 4 buildings from the person
num_simulations <- 10000   # Number of simulations

# Monte Carlo simulation
set.seed(123)  # For reproducibility
successes <- replicate(num_simulations, {
  # Randomly select 10 buildings from the 150
  selected_buildings <- sample(c(rep(1, person_buildings), rep(0, total_buildings - person_buildings)), top10_buildings)
  # Count how many of the selected buildings belong to the person
  sum(selected_buildings)
})
successes
# Estimate the probability of getting 4 or more of the person's buildings in the top 10
p_value <- mean(successes >= success_threshold)
# Print result
cat("Estimated probability (p-value) of getting 4 or more of the person's buildings in the top 10:", p_value, "\n")
```

## Problem 3: Permutation testing <small>(8pts)</small>

Below are data arising from a (fictionalized) data source: the number of defects per day on an assembly line before and after installation of a new torque converter (this is a totally fictional "part" of an assembly line--just treat these as "control" and "treatment" groups, respectively).

```{r}
before = c(4,5,6,3,6,3,4,5,5,3,4,6,4,6,3,4,2,2,0,7,5,8,4,5,1,4,4,8,2,3)
after  = c(3,2,4,3,7,5,5,2,2,4,5,2,2,6,1,5,6,3,2,3,7,3,4,5,4,2,2,6,7,8)
mean(before)
mean(after)
#t test no imame zavisen primerok treba da se testira razlikata vo odnos na nula
test_results <- t.test(before, after, var.equal = TRUE)
test_results
test_results <- t.test(before, after, var.equal = FALSE)
test_results
diference<-before-after

t_diference<-t.test(diference, mu = 0, alternative = "two.sided")
t_diference

```

a) (4 points) Use a permutation test to assess the claim that installation of the new part *changed* the prevalence of defects. That is, test the null hypothesis that the distribution of defects is the same before and after installation of the new part. Produce a p-value and interpret the results of your test in context.

```{r}

# Observed difference in means
obs_diff <- mean(before) - mean(after)
# Combine both groups
combined <- c(before, after)
n_before <- length(before)
n_after <- length(after)

# Set number of permutations
num_permutations <- 10000

# Permutation test
set.seed(123)  # For reproducibility
perm_diffs <- replicate(num_permutations, {
  permuted <- sample(combined)  # Permute the combined dataset
  perm_before <- permuted[1:n_before]  # Assign the first part as before
  perm_after <- permuted[(n_before + 1):(n_before + n_after)]  # Assign the second part as after
  mean(perm_before) - mean(perm_after)  # Calculate difference in means
})

# Calculate p-value (two-sided test)
p_value <- mean(abs(perm_diffs) >= abs(obs_diff))

# Print results
cat("Observed difference in means:", obs_diff, "\n")
cat("p-value from permutation test:", p_value, "\n")
```

b) (4 points) Explain, briefly, what you did above and why. Imagine that you are trying to explain to someone who isn't well versed in statistics what exactly you are doing in a permutation test. Explain your conclusion based on your test above. Three to five sentences should be plenty, but you are free to write as much or as little as you think is necessary to clearly explain your findings.



## Problem 4: Memes <small>(8pts)</small>

The following question comes from Karl Rohe, who developed the very first version of this class. This question has been reproduced in nearly the exact original (very amusing) wording.

> **Memes, part 1** (Please forgive me. I drank too much coffee before writing this question.)
> 
> In class thus far, there have been 416 comments posted in the bbcollaborate chat during class. An expert panel has judged 47 of these comments to be memes. The big-bad-deans say that they are concerned "if there is evidence that more than 10% of comments are memes." So, this looks like bad news, 47/416>10%.
> 
> Karl pleads with the deans: "Please, oh please, you big-bad-deans... Memeing is totally random." (I don't actually know what this notion of "random" means, but please just run with it for this question.) Then, along comes you, a trusty and dedicated 340 student. You say that "because we have only observed 416 comments, we don't really know what the 'true proportion' of memes."
> 
> 4a: What would be a good distribution for the number of memes?

> 
> 4b: Using your distribution from 4a, test the null hypothesis that the 'true proportion' is actually 10%. It's all up to you now... report the p-value.

Hints:

4a:  There should be a (hopefully) fairly intuitive choice of random variable that makes sense here. Look at your list of random variables and ask yourself which of these makes the most sense.

This is the test for proportions. Question is whether proportion of 47 judged memes among the 416 comments is greater than p0=10%.

H0: p=0.10   
HA: p>0.10

Random variable proportion p has binomial distribution. With sample n= 416 and p=47/416.
we can conclude that n is large enough to approximate binomial distribution with normal distribution:   np>=5  and n(1-p)>=5   

p0=0.1 and SD=SQRT(p0*(1-p0)/n)=0.0147
z=(p-p0)/SD
z=(47/416-0.1)/SQRT(0.1*(1-0.1)/416)~0.88

- For 4b, you can use the built-in function in R to simulate observations according to your null. Remember that you **always simulate *assuming* the null hypothesis**. Make sure your choice of the necessary parameter(s) reflects this assumption.


```{r}
p=1-pnorm(0.88,0,1)
p

```
We conclude that there is no enough evidence to reject H0. So among comments 10% are memes
